Modeling the Circumnuclear Mediums and Accetion Rates of Quiescent Supermassive Black Holes

### 2015-01-29 ms.tex
Minor adjustmments to equations and table.

\begin{table*}
\centering
\begin{minipage}{18cm}
\caption{Summary of Numerical Solutions}
\begin{tabular}{lccccccccccccc}
\hline
{$M_{\bullet,8}$} & {$v_{w}$} & {$\Gamma^{(a)}$} & {$r_b$} & {$\dot{M}/\dot{M}_{\rm edd}$} & {$(t_{\rm cool}/t_{\rm ff})_{r_s}$} & {comments}\\
\hline
0.01 & 300 & 0.8 & N/A & $ 3.2 \times 10^{ -4 }$ & 4.7 & rb in fact
unresolved \\
0.01 & 600 & 0.8 & N/A & $ 5.5 \times 10^{ -5 }$ & $ 3.2 \times
10^{ 2 }$ &  rb in fact unresolved\\
0.01 & 1200 & 0.8 & N/A & $ 9.4 \times 10^{ -6 }$ & $ 1.5 \times
10^{ 4 }$ &  rb in fact unresolved\\
0.1 & 300 & 0.8 & 100.0 & $ 2.5 \times 10^{ -3 }$ & 0.29 & energy
cons on outer grid \\
0.1 & 600 & 0.8 & 100.0 & $ 1.8 \times 10^{ -4 }$ & 97 &  \\
0.1 & 1200 & 0.8 & 100.0 & $ 3.1 \times 10^{ -5 }$ & $ 4.8 \times
10^{ 3 }$ &  energy cons on inner grid\\
1 & 300 & 0.8 & 100.0 & $ 1.6 \times 10^{ -2 }$ & $ 5.6 \times 10^{
  -2 }$ &  TI\\
1 & 600 & 0.8 & 100.0 & $ 7.3 \times 10^{ -4 }$ & 21 &  \\
1 & 1200 & 0.8 & 100.0 & $ 9.9 \times 10^{ -5 }$ & $ 1.5 \times 10^{
  3 }$ &  \\
1 & 300 & 0.8 & 50.0 & $ 8.2 \times 10^{ -3 }$ & 0.19 &  \\
1 & 300 & 0.8 & 200.0 & $ 3.6 \times 10^{ -2 }$ & $ 1.7 \times 10^{
  -2 }$ &  TI/convergence \\
%1 & 300.0 & 0.8 & 300.0 & $ 5.2 \times 10^{ -2 }$ & $ 9.8 \times 10^{
%  -3 }$ &  TI/convergence\\
1 & 300 & 0.8 & 400.0 &  &  &  TI/run unsuccessful\\
0.1 & 300 & 0.8 & 50.0 & $ 2.0 \times 10^{ -3 }$ & 0.46 &  \\
0.1 & 300 & 0.8 & 200.0 & $ 5.3 \times 10^{ -3 }$ & 0.12 &  TI/convergence\\
%0.1 & 300.0 & 0.8 & 300.0 & $ 7.2 \times 10^{ -3 }$ & $ 6.7 \times
%10^{ -2 }$   & TI/convergence  \\
0.1 & 300 & 0.8 & 400.0 & $ 8.9 \times 10^{ -3 }$ & $ 4.7 \times
10^{ -2 }$ & TI/convergence \\
0.01 & 1200 & 0.1 & N/A & $ 1.5 \times 10^{ -6 }$ & $ 4.1 \times
10^{ 4 }$ &  rb is unsresolved\\
0.1 & 1200 & 0.1 & N/A & $ 9.5 \times 10^{ -6 }$ & $ 5.7 \times 10^{
  3 }$ &  tried run in which rb=100 pc was resolved what would happen if rb=400 pc?\\
1 & 1200 & 0.1 & 100.0 & $ 6.0 \times 10^{ -5 }$ & $ 8.6 \times 10^{ 2 }$ &  \\
1 & 1200 & 0.1 & 300.0 & $ 1.9 \times 10^{ -4 }$ & $ 1.3 \times 10^{
  2 }$ &  \\
0.01 & 1200 & 0.1 & 100.0 & $ 1.5 \times 10^{ -6 }$ & $ 4.1 \times 10^{ 4 }$ &  \\
0.1 & 1200& 0.1 & 100.0 & $ 9.5 \times 10^{ -6 }$ & $ 5.7 \times 10^{ 3 }$ &  \\
1 & 1200 & 0.1 & 100.0 & $ 6.0 \times 10^{ -5 }$ & $ 8.6 \times 10^{ 2 }$ &  \\
1 & 1200 & 0.1 & 50.0 & $ 4.9 \times 10^{ -5 }$ & $ 1.2 \times 10^{ 3 }$ &  \\
1 & 1200 & 0.1 & 200.0 & $ 1.0 \times 10^{ -4 }$ & $ 3.4 \times 10^{ 2 }$ &  \\
%1 & 1200.0 & 0.1 & 300.0 & $ 1.9 \times 10^{ -4 }$ & $ 1.3 \times 10^{ 2 }$ &  convergence\\
1 & 1200 & 0.1 & 400.0 & $ 1.8 \times 10^{ -2 }$ & 0.37 &  TI/convergence\\
\hline
\label{table:models}  
\end{tabular}
$^{(a)}$ Inner power-law index of stellar Nuker profile.  $^{(b)}$ Break radius of stellar Nuker profile.
\end{minipage}
\end{table*}

@@

What is the effect of the Ia radius?  Between subsequent supernovae, stars release a gaseous mass $M_{\rm g} \approx \eta M_{\star}\tau_{\rm Ia}/t_{\rm h}$ interior to the Ia radius, which is gravitationally bound to the SMBH by an energy $E_{\rm bind} \sim M_{\rm g}\sigma^{2}$.  Thus, according to the above definitions, $E_{\rm Ia}/E_{\rm bind} \sim (v_{w}^{\rm Ia})^{2}/2\sigma^{2} \gtrsim 1$.  Hence Ia SN can in principle dynamically clear out gas from radii $\sim r_{\rm Ia} \gtrsim r_{\rm s}$.  Thus, even in cases when $r_{\rm s} > r_{\rm Ia}$, the SMBH accretion rate is limited to the value set by substituting the Ia radius $r_{\rm Ia}$ (eq.~[\ref{eq:rIa}]) for $r_{\rm s}$ in the derivation of equation $\ref{eq:eddr_analytic}$, i.e.  
\begin{eqnarray}
\left.\frac{\dot{M}}{\dot{M}_{\rm edd}}\right|_{\rm max} &\approx& \frac{\eta M_{\bullet}}{\dot{M}_{\rm edd} t_{\rm h}}\left(\frac{R_{\rm Ia}}{R_{\rm soi}}\right)^{2-\Gamma} \approx \nonumber \\
 && \begin{cases}
    3 \times 10^{-4} M_{\bullet,8}^{-1.26}(\tau_{\star}/t_{\rm h})^{-0.3}
   & \text{core} \\
    1.5 \times 10^{-4} \Mbheight^{-0.77}(\tau_{\star}/t_{\rm h})^{-0.7}   & \text{cusp}.
  \end{cases}
  \label{eq:eddr_Ia}
\end{eqnarray}
{\bf BDM: check!}

@@

{\bf AG: Note that I obtain Compton heating which is a factor of 2
  higher--although power law dependences look right to me. Note that
  the Compton heating will depend on radius--I use rs in my own
  estimate.  BDM: if so, the please update the numbers!  }
The importance of Compton heating can be estimated by assuming the SMBH
radiates with a luminosity $L = \epsilon_{\rm rad} \dot{M}c^{2}$, where
$\epsilon_{\rm rad}$ is the radiative efficiency and where $\dot{M}$ is
estimated from equation (\ref{eq:mdot_analytic}).  Then using
equations (\ref{eq:rs_simple}), (\ref{eq:mdot_analytic}),
(\ref{eq:rhors}), (\ref{eq:rhostarrs}) we calculate that from equation
(\ref{eq:vw_eff}) that the effective Compton heating rate at the stagnation radius is given by
\begin{align} v_{w}^{\rm C} \simeq
  \begin{cases} 13 \eta_{0.02}^{0.5}T_{\rm
C,9}^{0.5}\epsilon_{-2}^{0.5} M_{\bullet,8}^{0.38}v_{500}^{-1.4}\,{\rm
km\,s^{-1}} &, \text{core}\\ 20 \eta_{0.02}^{0.5}T_{\rm
C,9}^{0.5}\epsilon_{-2}^{0.5} M_{\bullet,8}^{0.24}v_{500}^{-0.7}\,{\rm
km\,s^{-1}} &, \text{cusp},
  \end{cases}
  \label{eq:vC}
\end{align} where $T_{C,9} = T_{C}/10^{9}$ K and $\epsilon_{-2} =
\epsilon_{\rm rad}/0.01 \sim 1$ is the maximum radiative efficiency for an advective
accretion flow expected to produce high $T_{\rm C}$ emission (e.g.~\citealt{Narayan&Yi95}).  If heating is provided by Compton heating alone, i.e. $\tilde{v}_{w} = v_{w}^{C}$, then equation (\ref{eq:vC}) can be solved for $\tilde{v}_{\rm w}$ to find
\begin{align} v_{w}^{\rm C} \simeq
  \begin{cases} 106 \eta_{0.02}^{0.21}T_{\rm
C,9}^{0.21}\epsilon_{-2}^{0.21} M_{\bullet,8}^{0.16}\,{\rm km\,s^{-1}}
&, \text{core}\\ 76 \eta_{0.02}^{0.29}T_{\rm
C,9}^{0.29}\epsilon_{-2}^{0.29} M_{\bullet,8}^{0.14}\,{\rm km\,s^{-1}}
&, \text{cusp},
  \end{cases}
  \label{eq:vC2}
\end{align} 
The fact that $v_{w}^{\rm C} \lesssim 200$ km s$^{-1}$ for physically realized values of $\eta \lesssim 1$ shows that Compton heating alone is insufficient to stabilize the flow against cooling
instability, i.e. $v_{w}^{C} < v_{\rm CI}$ (eq.~[\ref{eq:cooling2}]).

>>>>

\begin{table*}
\centering
\begin{minipage}{18cm}
\caption{Summary of Numerical Solutions}
\begin{tabular}{lccccccccccccc}
\hline
{$M_{\bullet,8}$} & {$v_{w}$} & {$\Gamma^{(a)}$} & {$r_b$} & {$\dot{M}/\dot{M}_{\rm edd}$} & {$(t_{\rm cool}/t_{\rm ff})_{r_s}$} & {comments}\\
\hline
0.01 & 300 & 0.8 & N/A & $ 3.2 \times 10^{ -4 }$ & 4.7 & rb in fact
unresolved \\
0.01 & 600 & 0.8 & N/A & $ 5.5 \times 10^{ -5 }$ & $ 3.2 \times
10^{ 2 }$ &  rb in fact unresolved\\
0.01 & 1200 & 0.8 & N/A & $ 9.4 \times 10^{ -6 }$ & $ 1.5 \times
10^{ 4 }$ &  rb in fact unresolved\\
0.1 & 300 & 0.8 & 100.0 & $ 2.5 \times 10^{ -3 }$ & 0.29 & energy
cons on outer grid \\
0.1 & 600 & 0.8 & 100.0 & $ 1.8 \times 10^{ -4 }$ & 97 &  \\
0.1 & 1200 & 0.8 & 100.0 & $ 3.1 \times 10^{ -5 }$ & $ 4.8 \times
10^{ 3 }$ &  \\
1 & 300 & 0.8 & 100.0 & $ 1.6 \times 10^{ -2 }$ & $ 5.6 \times 10^{
  -2 }$ &  TI\\
1 & 600 & 0.8 & 100.0 & $ 7.3 \times 10^{ -4 }$ & 21 &  \\
1 & 1200 & 0.8 & 100.0 & $ 9.9 \times 10^{ -5 }$ & $ 1.5 \times 10^{
  3 }$ &  \\
1 & 300 & 0.8 & 50.0 & $ 8.2 \times 10^{ -3 }$ & 0.19 &  \\
1 & 300 & 0.8 & 200.0 & $ 3.6 \times 10^{ -2 }$ & $ 1.7 \times 10^{
  -2 }$ &  TI/convergence \\
%1 & 300.0 & 0.8 & 300.0 & $ 5.2 \times 10^{ -2 }$ & $ 9.8 \times 10^{
%  -3 }$ &  TI/convergence\\
1 & 300 & 0.8 & 400.0 &  &  &  TI/run unsuccessful\\
0.1 & 300 & 0.8 & 50.0 & $ 2.0 \times 10^{ -3 }$ & 0.46 &  \\
0.1 & 300 & 0.8 & 200.0 & $ 5.3 \times 10^{ -3 }$ & 0.12 &  TI/convergence\\
%0.1 & 300.0 & 0.8 & 300.0 & $ 7.2 \times 10^{ -3 }$ & $ 6.7 \times
%10^{ -2 }$   & TI/convergence  \\
0.1 & 300 & 0.8 & 400.0 & $ 8.9 \times 10^{ -3 }$ & $ 4.7 \times
10^{ -2 }$ & TI/convergence \\
0.01 & 1200 & 0.1 & N/A & $ 1.5 \times 10^{ -6 }$ & $ 4.1 \times
10^{ 4 }$ &  rb is unsresolved\\
0.1 & 1200 & 0.1 & N/A & $ 9.5 \times 10^{ -6 }$ & $ 5.7 \times 10^{
  3 }$ &  tried run in which rb=100 pc was resolved what would happen if rb=400 pc?\\
0.01 & 1200 & 0.1 & 100.0 & $ 1.5 \times 10^{ -6 }$ & $ 4.1 \times 10^{ 4 }$ &  \\
0.1 & 1200& 0.1 & 100.0 & $ 9.5 \times 10^{ -6 }$ & $ 5.7 \times 10^{ 3 }$ &  \\
1 & 1200 & 0.1 & 100.0 & $ 6.0 \times 10^{ -5 }$ & $ 8.6 \times 10^{ 2 }$ &  \\
1 & 1200 & 0.1 & 50.0 & $ 4.9 \times 10^{ -5 }$ & $ 1.2 \times 10^{ 3 }$ &  \\
1 & 1200 & 0.1 & 200.0 & $ 1.0 \times 10^{ -4 }$ & $ 3.4 \times 10^{ 2 }$ &  \\
%1 & 1200.0 & 0.1 & 300.0 & $ 1.9 \times 10^{ -4 }$ & $ 1.3 \times 10^{ 2 }$ &  convergence\\
1 & 1200 & 0.1 & 400.0 & $ 1.8 \times 10^{ -2 }$ & 0.37 &  TI/convergence\\
\hline
\label{table:models}  
\end{tabular}
$^{(a)}$ Inner power-law index of stellar Nuker profile.  $^{(b)}$ Break radius of stellar Nuker profile.
\end{minipage}
\end{table*}

@@

What is the effect of the Ia radius?  Between subsequent supernovae, stars release a gaseous mass $M_{\rm g} \approx \eta M_{\star}\tau_{\rm Ia}/t_{\rm h}$ interior to the Ia radius, which is gravitationally bound to the SMBH by an energy $E_{\rm bind} \sim M_{\rm g}\sigma^{2}$.  Thus, according to the above definitions, $E_{\rm Ia}/E_{\rm bind} \sim (v_{w}^{\rm Ia})^{2}/2\sigma^{2} \gtrsim 1$.  Hence Ia SN can in principle dynamically clear out gas from radii $\sim r_{\rm Ia} \gtrsim r_{\rm s}$.  Thus, even in cases when $r_{\rm s} > r_{\rm Ia}$, the SMBH accretion rate is limited to the value set by substituting the Ia radius $r_{\rm Ia}$ (eq.~[\ref{eq:rIa}]) for $r_{\rm s}$ in the derivation of equation $\ref{eq:eddr_analytic}$, i.e.  
\begin{eqnarray}
\left.\frac{\dot{M}}{\dot{M}_{\rm edd}}\right|_{\rm max} &\approx& \frac{\eta M_{\bullet}}{\dot{M}_{\rm edd} t_{\rm h}}\left(\frac{R_{\rm Ia}}{R_{\rm soi}}\right)^{2-\Gamma} \approx \nonumber \\
 && \begin{cases}
    4.5 \times 10^{-4} M_{\bullet,8}^{-1.33}(\tau_{\star}/t_{\rm h})^{1.1}
   & \text{core} \\
    2.2 \times 10^{-4} \Mbheight^{-0.84}(\tau_{\star}/t_{\rm h})^{0.7}   & \text{cusp}.
  \end{cases}
  \label{eq:eddr_Ia}
\end{eqnarray}
{\bf AG: Fixed this expression. I will check again! Ia accretion rate should
  increase with time, as in the burst limit the Ia rate would decline
  with time, so the Ia radius would move outwards (need more volume to
  get a supernova per dynamical time.}

@@

{\bf AG: Note that the Compton heating will depend on the distance of
  the hole as $1/r^2$ so Compton heating will be stronger closer to the
  black hole! }
The importance of Compton heating can be estimated by assuming the SMBH
radiates with a luminosity $L = \epsilon_{\rm rad} \dot{M}c^{2}$, where
$\epsilon_{\rm rad}$ is the radiative efficiency and where $\dot{M}$ is
estimated from equation (\ref{eq:mdot_analytic}).  Then using
equations (\ref{eq:rs_simple}), (\ref{eq:mdot_analytic}),
(\ref{eq:rhors}), (\ref{eq:rhostarrs}) we calculate that from equation
(\ref{eq:vw_eff}) that the effective Compton heating rate at the stagnation radius is given by
\begin{align} v_{w}^{\rm C} \simeq
  \begin{cases} 29 \eta_{0.02}^{0.5}T_{\rm
C,9}^{0.5}\epsilon_{-2}^{0.5} M_{\bullet,8}^{0.38}v_{500}^{-1.4}\,{\rm
km\,s^{-1}} &, \text{core}\\ 50 \eta_{0.02}^{0.5}T_{\rm
C,9}^{0.5}\epsilon_{-2}^{0.5} M_{\bullet,8}^{0.24}v_{500}^{-0.7}\,{\rm
km\,s^{-1}} &, \text{cusp},
  \end{cases}
  \label{eq:vC}
\end{align} where $T_{C,9} = T_{C}/10^{9}$ K and $\epsilon_{-2} =
\epsilon_{\rm rad}/0.01 \sim 1$ is the maximum radiative efficiency
for an advective accretion flow expected to produce high $T_{\rm C}$
emission (e.g.~\citealt{Narayan&Yi95}).  If heating is provided by
Compton heating alone, i.e. $\tilde{v}_{w} = v_{w}^{C}$, then equation
(\ref{eq:vC}) can be solved for $\tilde{v}_{\rm w}$ to find
\begin{align} v_{w}^{\rm C} \simeq
  \begin{cases} 106 \eta_{0.02}^{0.21}T_{\rm
C,9}^{0.21}\epsilon_{-2}^{0.21} M_{\bullet,8}^{0.16}\,{\rm km\,s^{-1}}
&, \text{core}\\ 76 \eta_{0.02}^{0.29}T_{\rm
C,9}^{0.29}\epsilon_{-2}^{0.29} M_{\bullet,8}^{0.14}\,{\rm km\,s^{-1}}
&, \text{cusp},
  \end{cases}
  \label{eq:vC2}
\end{align} 
The fact that $v_{w}^{\rm C} \lesssim 200$ km s$^{-1}$ for physically realized values of $\eta \lesssim 1$ shows that Compton heating alone is insufficient to stabilize the flow against cooling
instability, i.e. $v_{w}^{C} < v_{\rm CI}$ (eq.~[\ref{eq:cooling2}]).

### 2015-01-29 windheat.tex
Minor adjustmments to equations and table.

The quantity of mass lost in PMS winds $\Delta M(t)$ is estimated from the expression given by \citet{CiottiOstriker:2007a} (their eq.~[10]),
\begin{align}
\Delta M=
\begin{cases}
0.945 M_{\rm TO}-0.503 & M_{\rm TO} < 9 \Msun\\
 M_{\rm TO}-1.4 \Msun &  M_{\rm TO} \ge 9 \Msun,
\end{cases}
\end{align}
where $M_{\rm TO}$ is the turn-off mass, which at time $t$ is calculated as
\begin{align}
\log(M_{\rm TO})/M_{\odot} =0.24 + 0.068 x^2-0.34 x+4.76 e^{-4.58 x},
\end{align}
where $x=\log(t/10^9 {\rm years})$.  This functional fit is designed to reproduce the results of \citet{MaederMeynet:1987a} (their Table 9) for massive stars while asymptoting to the formula provided by \citet{CiottiOstriker:2007a} (their eq.~[9]) for intermediate and late times ($t \gsim 10^8$ years).

@@

The effective wind velocity in the limit of an impulsive star formation may then be written
as 
\begin{align}
\bar{v}_w(t)=2 \dot{\bar{e}}(t)/\dot{\bar{m}}(t)
\label{eq:vwImp}
\end{align}
while {\bf BDM: insert correct expression}
\begin{align}
\eta = \dot{\bar{m}}(t)
\label{eq:vwImp}
\end{align}
We plot $\bar{v}_w(t)$ and $\eta(t)$ in Figures~\ref{fig:vwImp} and~\ref{fig:etaImp}.

>>>>

The quantity of mass lost in PMS winds $\Delta M(t)$ is estimated from the expression given by \citet{CiottiOstriker:2007a} (their eq.~[10]),
\begin{align}
\Delta M=
\begin{cases}
0.945 M_{\rm TO}-0.503 & M_{\rm TO} < 9 \Msun\\
 M_{\rm TO}-1.4 \Msun &  M_{\rm TO} \ge 9 \Msun,
\end{cases}
\end{align}
where $M_{\rm TO}$ is the turn-off mass, which at time $t$ is calculated as
\begin{align}
\log(M_{\rm TO})/M_{\odot} =0.24 + 0.068 x^2-0.34 x+4.76 e^{-4.58 x},
\end{align}
where $x=\log(t/10^9 {\rm years})$.  This functional fit is designed
to reproduce the results of \citet{MaederMeynet:1987a} (their Table 9)
for massive stars while asymptoting to the formula provided by
\citet{CiottiOstriker:2007a} (their eq.~[9]) for intermediate and late
times ($t \gsim 10^8$ years). {\bf AG: Mathematical aside--this
  expression divereges for t>>$t_{hubble}$}

@@

The effective wind velocity in the limit of an impulsive star formation may then be written
as 
\begin{align}
\bar{v}_w(t)=2 \dot{\bar{e}}(t)/\dot{\bar{m}}(t)
\label{eq:vwImp}
\end{align}
while {\bf AG: eta expression is now fixed--note that
  $\dot{\bar{m}}(t)$ is the mass injection rate per stellar mass, so
  that eta is dimensionless as it should be }
\begin{align}
\eta = \dot{\bar{m}}(t) \th
\label{eq:vwImp}
\end{align}
We plot $\bar{v}_w(t)$ and $\eta(t)$ in Figures~\ref{fig:vwImp} and~\ref{fig:etaImp}.
